Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-05-16 PAPER_JA_edits.tex
Wed May 16 10:08:39 EDT 2018

We can again calculate the estimator given by \ref{eq:} explicitly by using the Sherman-Morrison identity to invert the covariance matrix.  Defining
\begin{equation}
\xi \equiv \frac{u/s}{1+u/s}
\end{equation} 
we have
\begin{equation}
\invC  =   \frac{1}{s} \left(\I - \xi \mathbf{m m}^T \right)
\end{equation}
and the virtue of weighting by $\invC$ becomes clearer: in the limit $u \gg s$, $\xi \to 1$, and this becomes
\begin{equation}
\invC  = \frac{1}{s} \left(\I - \mathbf{m m}^T \right)
\end{equation}
where $\I - \mathbf{m m}^T$ is the projection operator, projecting out $\mathbf{m}$ from any any vector it acts on.   Proceeding to calculate the full estimator
\begin{equation}
\hat{s} = \half {\F}^{-1} \left(\bm \Delta^T \invC \PDeriv{\C}{s}  \invC \bm \Delta \right) = \frac{ \bm \Delta^T  (\I +  (\xi^2 - 2 \xi)  \mathbf{m m}^T)  \bm \Delta}{N + \xi^2 - 2 \xi} 
\end{equation}
we note that the expectation is
\begin{equation}
\langle \hat{s} \rangle = \frac{ N(s + u + (\xi^2 - 2 \xi) 
[u + \langle(\bm{\sigma} \cdot \mathbf{m})^2 \rangle])} 
{N + \xi^2 - 2 \xi} 
\end{equation}
This again has a simple interpretation in the limit $\xi \to 1$, where
\begin{equation}
\langle \hat{s} \rangle = \frac{ s - \langle(\bm{\sigma} \cdot \mathbf{m})^2\rangle} {N -1} 
\end{equation}

>>>>

We can again calculate the estimator given by \ref{eq:} explicitly by using the Sherman-Morrison identity to invert the covariance matrix.  Defining
\begin{equation}
\xi \equiv \frac{u/s}{1+u/s}
\end{equation} 
we have
\begin{equation}
\invC  =   \frac{1}{s} \left(\I - \xi \mathbf{m m}^T \right)
\end{equation}
and the virtue of weighting by $\invC$ becomes clearer: in the limit $u \gg s$, $\xi \to 1$, and this becomes
\begin{equation}
\invC  = \frac{1}{s} \left(\I - \mathbf{m m}^T \right)
\end{equation}
where $\I - \mathbf{m m}^T$ is the projection operator, projecting out $\mathbf{m}$ from any any vector it acts on.   Proceeding to calculate the full estimator
\begin{equation}
\hat{s} = \half {\F}^{-1} \left(\bm \Delta^T \invC \PDeriv{\C}{s}  \invC \bm \Delta \right) = \frac{ \bm \Delta^T  (\I +  (\xi^2 - 2 \xi)  \mathbf{m m}^T)  \bm \Delta}{N + \xi^2 - 2 \xi} 
\end{equation}
The expectation is
\begin{equation}
\langle \hat{s} \rangle = \frac{ N(s + u + (\xi^2 - 2 \xi) 
[u + \langle(\bm{\sigma} \cdot \mathbf{m})^2 \rangle])} 
{N + \xi^2 - 2 \xi} 
\end{equation}
This again has a simple interpretation in the limit $\xi \to 1$, where
\begin{equation}
\langle \hat{s} \rangle = \frac{N}{N -1}  (s - \langle(\bm{\sigma} \cdot \mathbf{m})^2\rangle )
\end{equation}
Thus, in the limit the contaminant is large compared to $s$, and to the extent that $\bm \sigma \cdot \mathbf{m} \approx 0$, this yields a nearly unbiased estimator for $s$ and the inverse covariance weighting has completely removed the effect of the contaminant.
